---
title: "Example plot"
author: "Mamoon Y. Ismail"
date: "`r Sys.Date()`"
output: html_document
---

**BOLD**

_ITALICS_

1- The backticks (`) indicate code

2- r indicates R code


3- In-line code: is how to include r code results into text

The version of R is `r getRversion()`

4- It’s usually best to give each code chunk a name, like chunk_name ```{r chunk_name}

5- echo=FALSE indicates that the code will not be shown in the final document like this ```{r chunk_name, echo=FALSE}

6- Use results="hide" to hide the results/output (but here the code would still be displayed), ```{r chunk_name, results="hide"}

7- Use include=FALSE to have the chunk evaluated, but neither the code nor its output displayed, ```{r chunk_name, include=FALSE}

8- For figures, you’ll want to use options like fig.width and fig.height. For example: {r scatterplot, fig.width=8, fig.height=6}. Note that if include=FALSE, all of the code, results, and figures will be suppressed. If include=TRUE and results="hide", the results will be hidden but figures will still be shown. To hide the figures, use fig.show="hide". NOTE: results=hide is wrong; you need results="hide".

9- Global chunk options: use the same set of chunk options throughout a document to avoid retype it in every chunk, so you want to set some global chunk options at the top of your document. For example, I might use include=FALSE or at least echo=FALSE globally for a report to a scientific collaborator who wouldn’t want to see all of the code. And I might want something like fig.width=12 and fig.height=6 if I generally want those sizes for my figures.

10- If you want a particular chunk to have a different behavior, for example, to have a different figure height, you’d specify a different option within that chunk. For example: {r a_taller_figure, fig.height=32}.


### Loading librarry

A chunk needed to load the library

```{r loadlib, echo=TRUE}
library(tidyverse)
```

## Global Options


```{r global_options, include=TRUE}
#These are default settings eg fig size or whether you print code
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE,
                      fig.width=12, fig.height=8, fig.path='figs/')
```

## SURVEYS

This is a survey data that has been collected about animal species.



### Reading Data



Reading Colected data into a file called survey



```{r loaddata, echo=TRUE}
surveys <- read_csv("data/portal_data_joined.csv")
str(surveys)

```


## Plotting

Add plots of different relations

```{r Plot, echo=TRUE}
plot(as.factor(surveys$plot_id))
```



## Session Info:

Include “session info” in your document, preferably at the bottom: that lists the version of R that you’re using plus all of the packages you’ve loaded



```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```



