---
title: "R Homework Assignment"
author: "Mamoon Y. Ismail, Jeeban Panthi, Tiare Fridrich"
date: "`r Sys.Date()`"
output: html_document
---

**BOLD**

_ITALICS_

1- The backticks (`) indicate code

2- r indicates R code


3- In-line code: is how to include r code results into text

The version of R is `r getRversion()`

4- It’s usually best to give each code chunk a name, like chunk_name ```{r chunk_name}

5- echo=FALSE indicates that the code will not be shown in the final document like this ```{r chunk_name, echo=FALSE}

6- Use results="hide" to hide the results/output (but here the code would still be displayed), ```{r chunk_name, results="hide"}

7- Use include=FALSE to have the chunk evaluated, but neither the code nor its output displayed, ```{r chunk_name, include=FALSE}

8- For figures, you’ll want to use options like fig.width and fig.height. For example: {r scatterplot, fig.width=8, fig.height=6}. Note that if include=FALSE, all of the code, results, and figures will be suppressed. If include=TRUE and results="hide", the results will be hidden but figures will still be shown. To hide the figures, use fig.show="hide". NOTE: results=hide is wrong; you need results="hide".

9- Global chunk options: use the same set of chunk options throughout a document to avoid retype it in every chunk, so you want to set some global chunk options at the top of your document. For example, I might use include=FALSE or at least echo=FALSE globally for a report to a scientific collaborator who wouldn’t want to see all of the code. And I might want something like fig.width=12 and fig.height=6 if I generally want those sizes for my figures.

10- If you want a particular chunk to have a different behavior, for example, to have a different figure height, you’d specify a different option within that chunk. For example: {r a_taller_figure, fig.height=32}.


### Loading librarry

A chunk needed to load the library

```{r loadlib, echo=TRUE}
library(tidyverse)
```

## Global Options


```{r global_options, include=TRUE}
#These are default settings eg fig size or whether you print code
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE,
                      fig.width=12, fig.height=8, fig.path='figs/')
```

## SURVEYS

This is a survey data that has been collected about animal species.



### Reading Data



Reading Colected data into a file called survey



```{r loaddata, echo=TRUE}
surveys <- read_csv("~/Desktop/rclass/homework/Assignment-3R/data_output/surveys_NAsrem.csv")

```


## Plotting

Add plots of different relations



## Session Info:

Include “session info” in your document, preferably at the bottom: that lists the version of R that you’re using plus all of the packages you’ve loaded

These are the packages we need to deal with

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(dplyr)
library(ggpubr)
library(Hmisc)
```

###1. Relationship between sex and hindfoot length
```{r}
ggplot(surveys, aes(x=sex, y=hindfoot_length))+geom_boxplot(alpha = 1, color='red')+xlab('Sex')+ylab('Length of hindfoot (inch)')+stat_summary(fun.y=mean, color="blue", geom="point",  shape=15, size=1,show_guide = T, show.legend = T)
```
Figure 1: Hindfoot length for male and female. The blue dot indicates the mean (average) hindfoot length. 

###2. Number of individuals of each of the species classfied by sex
```{r}
ggplot(surveys, aes(species_id))+geom_bar(aes(fill=sex))+ylab('Number of individuals recorded')+xlab('Species ID') 
```
Figure 2: Number of individuals of each of the species classfied by sex

#3. Correlationship between weight and hindfoot length
```{r}
ggscatter(surveys, x = "weight", y = "hindfoot_length",  add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "pearson",  xlab = "Hindfoot Length (inch)", ylab = "Weight (gram)")
```
Figure 3: Scatter plot showing the correlation between weight and hindfoot length of species

The p-value of the test is less than the significance level alpha = 0.05. We can conclude that weight and hindfoot length are significantly positively correlated with a correlation coefficient of 0.68 and p-value of 2.2*10^{-16}

#4.4.	Hindfoot length of different species
```{r}
ggplot(data = surveys, aes(x = species_id, y = hindfoot_length)) +  geom_boxplot(alpha=1, color='red')+xlab('Species ID')+ylab('Hindfoot length (inch)')+stat_summary(fun.y=mean, color="blue", geom="point",  shape=15, size=1,show_guide = T, show.legend = T)
```
Figure 4: Box plot of hindfoot length of different species. The blue dot indicates the average (mean) hindfoot length.

###5 Number of species found yearly
```{r species_count}
# Yearly counts
year_counts <- surveys_NAsrem %>%
  group_by(year, species_id) %>% tally()
# Plot
ggplot(year_counts, aes(x=year, y=n, color=species_id)) +
  geom_col()+facet_wrap(~species_id)

```
Figure 5: time series faceting plot showing the count of each species yearly.

###6 Number of species found in each plot type
```{r species_count_by_plot}
# plot counts
Plot_counts <- surveys_NAsrem %>% group_by(year, plot_type) %>% tally()
# plot
ggplot(Plot_counts, aes(x=year, y=n, color=plot_type)) +
  geom_col()+facet_wrap(~plot_type)

```
Figure 6: time series faceting plot showing the count of species observed in each plot type yearly.

###7 Males and females Average weight yearly
```{r Sex_average_weight}
# Species mean weight 
spe_meanweight <- surveys_NAsrem %>% group_by(year,species_id, sex) %>%
  summarize(mean_weight = mean(weight))
# plot
ggplot(spe_meanweight, aes(x=year, y=mean_weight, color=sex)) +
  geom_line()+facet_wrap(~species_id) + theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(axis.text.x = element_text(colour = "grey20", size = 10, angle = 90, hjust = 0.5, vjust = 0.5))
```
Figure 7: time series faceting plot showing yearly average weight of each male and female species.

###8 Count of species observed
```{r species count}
# I'm making few plots here which are slightly different and we may end up chose some of them or none
ggplot(surveys_NAsrem) + stat_count(aes(x = species_id, fill = species_id))

# Count of male and female species observed
ggplot(surveys_NAsrem) + stat_count(aes(x = species_id, fill = sex))

# Count of male and female species observed dodged
ggplot(surveys_NAsrem) + stat_count(aes(x = species_id, fill = sex), position = "dodge")

```
Figure 8: Bar graph showing number of each species obesrevd

###9 relationship between species weight and sex
```{r weight_sex}
ggplot(surveys_NAsrem) + geom_boxplot(aes(x = species_id, y = weight, fill = sex)) +
  coord_flip()

```
Figure 9: Box plot (axis fliped) showing the relationship between species sex and weight





